Cryptographic assurance of secrecy and fairness in voting

### 2012-12-05 paper/paper.tex
Rework descriptions, some analysis, and add topics
Reworked (largely rewrote) system descriptions for Scantegrity and
Helios.

Did some work (though not much) on the analysis -- in part, moving text
that I had elsewhere in the paper.

Added some new topics for discussion.

Also introduced distinction between verifiable recording and verifiable
collection (as discussed face-to-face and as pointed out in the Prêt à
Voter paper).

% title
\title{Cryptographic assurance of fairness in voting}
\date{}
\author{
\begin{tabular}{c c c}
Kelsey Francis & Curtis Free & Christopher Martin \\
\small \tt{francis@gatech.edu} & \small \tt{curtis.free@gatech.edu} & \small \tt{chris.martin@gatech.edu}
\end{tabular}
}

@@

When a voter casts a ballot, he/she knows which candidate(s) he/she \emph{intended} to vote for --
but once the ballot is cast, a system that provides \term{collected-cast-cast verifiability} allows
the voter to compare the recorded vote to what he/she intended to indicate on the ballot.

@@

@@

Once votes have been recorded, it is the responsibility of the state to tally votes for the various
candidates in a race and publish \emph{accurate} tallies. Generally, the electorate is reliant on
\emph{someone} (e.g., the government at some level) to provide these tallies. But it is much less
common for voters themselves to have the opportunity to verify that the votes were tallied
\emph{exactly} as they were cast.

@@

\section{Scantegrity \cite{scantegrity_ii}}

With Scantegrity (actually Scantegrity II, successor to the original Scantegrity system), each
ballot created is assigned a unique number. While this number can be printed in plain sight, it can
be encrypted such that someone with physical access to a ballot box cannot link a ballot to someone
with a receipt.

Furthermore, each candidate on the ticket -- presented in a consistent manner across tickets (i.e.
no randomization, though some can be used) -- is selected by filling in a ``bubble'' that reveals an
alphanumeric confirmation code: these codes are selected such that they are independent and
uniformly random. On a standard voter ballot, only the codes for candidates selected by the voter
are revealed. The body holding the election -- e.g., the government -- keeps a secret \term{Table P}
that contains all confirmation codes generated for all tickets.

An important point regarding Scantegrity is that it relies on a \emph{physical property} -- the
invisible ink used when printing confirmation code ``bubbles'' -- to combat vote coercion (i.e., no
one can force a voter to produce a particular receipt, and no one can take a receipt and ably pay
the person for having voted a certain way).

Scantegrity II does provide -- in their terms:
\begin{enumerate}
	\item
		\term{Voter verifiability} - the ability of a voter to ensure that his/her vote was included in
		the tally; and
	\item
		\term{Universal verifiability} - the ability of \emph{anyone} to verify that, given that some
		set of votes were cast, the candidate counts are correct.
\end{enumerate}

Unfortunately, in order to provide some of its guarantees, Scantegrity II does require a published
list of \emph{who} voted. Such a list could aid groups from penalizing those who do not -- or do --
vote.

At the polling place, a voter may choose to receive a second \term{audit ballot}. This ballot is
marked as such, and all confirmation codes are revealed. The importance of this second ballot will
be revealed below.

The group holding the election generates four tables:
\begin{enumerate}
	\item
		\term{Table P}, which is kept secret and which contains all confirmation codes printed on all
		ballots (along with the corresponding candidate);
	\item
		\term{Table Q}, which is published and which uses scanned information along with Table P to
		reveal -- on a per-ballot basis -- only those confirmation codes which the voter actually
		revealed (including \emph{all} codes for audit ballots).
	\item
		\term{Table R}, which provides a randomly-order but correct bijection from cells in Table Q to
		cells in Table S. After the election has concluded, rows in this table are flagged with a
		boolean ``true'' value if they correspond to codes (Table Q) revealed by the voter on
		\emph{non-audit} ballots. At that time, only one side of the mapping -- to Q or to S -- is
		revealed for non-audit ballots, and which side is revealed is select by a ``true random,
		publicly verifiable coin flip.''
	\item
		\term{Table S}, which provides per-candidate counts by recording in cells, via Table R mappings,
		those codes that were actually revealed on legitimate non-audit ballots.
\end{enumerate}

\todo{Include own image of the tables.}

\todo{Need a much more in-depth discussion of how Scantegrity works. Right now it's a play-by-play
from the source.}

@@

\section{Helios \cite{helios}}

Overall, Helios performs the following steps in an election:
\begin{enumerate}
	\item
		Encrypt individual votes as they're cast.
	\item
		Mix the votes for privacy.
		\begin{enumerate}
    	\item
				Re-encrypt the ballots (ElGamal).
    	\item
				Shuffle using a \term{mixnet}.
			\item
				Give cryptographic proof of a legitimate shuffle.
			\item
				Decrypt the votes.
			\item
				Give cryptographic proof of the decryptions.
		\end{enumerate}
	\item
		Tally the votes.
\end{enumerate}

Helios relies on auditing at two stages: (a) by voters when casting ballots and (b) by independent
auditors after votes have been tallied.

After entering one's candidate choices, but before a vote is actually cast, a voter is presented
with a (SHA-1) hash of his/her encrypted vote and can choose to ``audit'' the ballot, in which case
Helios will reveal all the information necessary (including ``randomness'') for the voter to verify
that the hash correctly represents the vote he/she indended to cast. A voter can continue to audit
ballots (due to randomness, different each time) before casting one. He/she is \emph{not} given the
information necessary to audit the ballot that is ultimately cast.

Once votes have been tallied at the end of the election, auditors can obtain all data from the
election -- except for the secret keys necessary for shuffling and encryption but (notably)
including the generated proofs -- to verify that the votes were counted correctly. These auditors
can post their verifications publicly and compare results with other auditors.

The ballot shuffle and decryption are proved using a technique called an \term{Honest-Verifier
Zero-Knowledge} proof. However, because the challenges in Helios are generated without human
involvement, a technique called Fiat-Shamir is used to obtain a hash (so $\code{0}$s and
$\code{1}$s) that are treated as the challenges.

\todo{Need a much more in-depth discussion of how Helios works. Right now it's a play-by-play from
the source.}

@@

@@

In both Scantegrity and Helios, the following information is revealed after the election:
\begin{itemize}
	\item
		\emph{Some} information about \emph{all} encrypted ballots, linked to voters.
	\item
		\emph{All} decrypted ballots, emph{not} linked to voters.
\end{itemize}

In Scantegrity, ``public'' version of a ballot is really just an encryption of the voter's
selections: it is an independent, uniformly random code chosen beforehand and corresponding to a
particular candidate for that ballot. Table P defines an explicit, ``keyless'' encryption
function -- and each ballot holds just enough information from Table P to encrypt a single voter's
responses.

@@

\todo{We can also talk about Scantegrity's Table P as a definition wherein the ballot number is the
key.}

@@

In Scantegrity, the ballot ``mix'' is predetermined, as it is defined by the generation of Table R.
When that table is constructed, ballots are randomly \todo{Really?} ordered. With the exception of
audit ballots, the information revealed that is ultimately revealed allows half of the ballots to be
traced only from Table Q to Table R; and the other half can only be traced from Table R to Table S.
Should Table P be compromised, one could complete all mappings and violate secrecy entirely -- or
simply decrypt encrypted votes with no attention giving to Table R. Should the mapping on
\emph{either} size of Table R (to votes in Table Q or tallies in Table S) be deterministic, privacy
would be compromised for half of all voters.

@@

@@

@@

Scantegrity prevents coercion via the manner it uses to ``encrypt'' ballots. Once cast, a ballot's
ciphertext is \emph{always public}. A voter can claim that he/she voted in a particular manner, but
without exposure of the data maintained in Table P -- Scantegrity's ``secret'' -- the voter has no
way to prove to a third party that the claim is true. This is

@@

@@

>>>>

% title
\title{Cryptographic assurance of secrecy and fairness in voting}
\date{}
\author{
\begin{tabular}{c c c}
Kelsey Francis & Curtis Free & Christopher Martin \\
\small \tt{francis@gatech.edu} & \small \tt{curtis.free@gatech.edu} & \small \tt{chris.martin@gatech.edu}
\end{tabular}
}

@@

When casting a ballot, a voter knows which candidate(s) he/she \emph{intended} to vote for -- but
once the ballot is cast, a system that supports verification of vote reporting allows the voter to
compare the recorded vote to what he/she intended to indicate on the ballot.

@@

\subsection{Verifiable collection}

After a ballot has been cast, and the voter has left the polling place, how can he/she be certain
that the ballot used in the final count is what he/she actually cast? (This could be a problem even
if the user has verified the \emph{recording} of the vote.)

It is possible -- as we shall see -- for a voting system to allow a voter to check that the ballot
he/she cast is exactly what was considered during vote counting.

@@

Once votes have been collected, it is the responsibility of the state to tally votes for the various
candidates in a race and publish \emph{accurate} tallies. Generally, the electorate is reliant on
\emph{someone} (e.g., the government at some level) to provide these tallies. But it is much less
common for voters themselves to have the opportunity to verify that the votes were tallied
\emph{exactly} as they were cast.

@@

\section{Scantegrity}

The Scantegrity system \cite{scantegrity_ii} (technically, we consider an updated version known as
Scantegrity II) is designed to provide verifiability atop existing in-person voting systems that
rely on fill-in ballots that are processed via optical scan.

\subsection{Before voting}

Before an election is held, the body holding the election (e.g., the government) must construct four
Tables that lie at the core of the Scantegrity system: Tables P, Q, R, and S.

Some selection of persons associated with the electing body use a secret-sharing scheme to protect a
secret pseudorandom number generator key. Table P lists every ballot that will be produced for the
election and specifies exactly what code (which we'll see later) is used for each candidate per
ballot. These codes are generated using the pseudorandom generator and \emph{must} be kept secret.

Table Q has one row per ballot, as in Table P, but the codes presented on that ballot are included
in randomly-ordered cells so that a particular code cannot be mapped to a particular candidate.
Once ballots have been collected, only the cell actually marked on each ballot is revealed. (Note
that the column shuffle must be per-row, as otherwise one could determine whether two voters
cast a vote for the same candidate.)

Table S contains final vote tallies. Each column represents a particular candidate, and there will
be one ``checked'' cell in each column for each ballot cast for that candidate.

Table R is the ``heart'' of the Scantegrity Tables: it maps cast ballots in Table Q to tally marks
in Table S. Each row in Table R maps a single cell in Q to a single cell in S, and the order of
these mappings is pseudorandom (though the mappings must obey the data specified in Table P for vote
correctness).

\todo{Include our own depiction of the tables in an image.}

\subsection{Voting}

A Scantegrity ballot consists of two portions: a special optical-scan ballot strip and a second
piece of paper that the voter will keep as a receipt. Each ballot is identified by a unique number,
which is printed on both portions. (While this nuber can be printed in plain sight, it can also be
encrypted on the optical-scan portion such that someone with physical access to a ballot box cannot
link a particular ballot to a voter with receipt in-hand.)

All candidates are listed on the optical-scan ballot, and the voter makes his/her choice by filling
in a ``bubble'' that reveals an alphanumeric confirmation code (as specified in the private Table
P), which the voter then records on the receipt portion. The ballot is sent through an optical
scanner, and the voter takes his/her receipt when leaving the polling location. It can later be
compared against published Table Q to ensure that the vote was recorded properly.

Additionally, during the voting process, a voter can ask a poll worker for an second ``audit''
ballot. This ballot is initially just like any other ballot: however, \emph{all} candidate codes are
revealed, and the ballot is marked as an audit ballot. An audit ballot is scanned, but the voter is
allowed to keep the ballot for later verification.

\subsection{After voting}

When a ballot has been received, the following takes place:
\begin{enumerate}
	\item
		The candidate code is revealed in Table Q. (If this is an audit ballot -- see below -- then all
		are revealed, and we'll perform the following steps for each code.)
	\item
		A ``check'' is placed on the row in Table R that maps the revealed Q value to some value in S.
	\item
		The Table R mapping is followed into Table S, and a ``check'' is added to a cell in Table R
		(to be counted as a vote for that row's candidate).
\end{enumerate}

Tables Q and S and published in their entirety. However, one member of each Q-to-S cell mapping pair
is hidden prior to its publication (except for audit ballots, for which both sides of the mapping
are revealed). This data is required for third-party verification.

This process will be discussed further in the analysis.

@@

\todo{Fill in.}

\section{Helios}

Helios \cite{helios} -- unlike Scantegrity and Pr\^{e}t \`{a} Voter -- is an \emph{online} voting
system that attempts to meet many of the same goals (namely, verifiability).

\subsection{Before voting}

At the heart of Helios is a single ElGamal keypair. The secret key is known only to the Helios
server, but the public key can (by its nature) be accessed in some way by all participants.

\subsection{Voting}

When a voter goes to cast a vote, his/her candidate selection is encrypted locally (i.e., on his/her
computer terminal, in the browser) using ElGamal. Prior to casting the ballot, the voter is
presented with a SHA-1 hash of his/her encrypted ballot. At this point, the voter may choose either
to submit the ballot or to ``audit'' it. In the former case, the ballot is cast and voting is
complete. But when a voter chooses to ``audit,'' he/she is presented with \emph{all} of the data
used to encrypt the vote. Namely, this includes the public key and the ``randomness'' that were used
by an implementation of ElGamal to encrypt the vote. The voter can then -- using a modified ElGamal
(one which allows randomness to be specified) -- perform the encryption and SHA-1 hash him/herself
to verify that the hash that \emph{would} have been cast actually does represent the ballot he/she
intended to cast.

\subsection{After voting}

Once all ballots have been collected by the Helios server, they are shuffled by a cryptographic
``mixnet.'' \todo{We should discuss re-encryption details, which are covered in the Helios paper.
Maybe we cover the details in the Pr\^{e}t \`{a} Voter discussion?}

The ``mixnet'' accomplishes two goals: it shuffles the votes \emph{and} re-encrypts them (with the
same key, but with different random inputs). A random permutation -- referred to as $\pi_{N}$ (where
$N$ is the total number of votes) -- is first used to perform the shuffle. Each ciphertext 
$c_{i}$ is then ``re-encrypted'' to a new ciphertext $d_{i}$ that equivalent to the first \emph{had
the random input at encrypt time been different} by a particular amount $s_{i}$.

The ``shuffling'' is important because it prevents an attacker (or simple onlooker) from linking
a decrypted ballot with a particular encrypted ballot -- and thus a voter. Re-encryption means that
an auditor examining a proof of proper decryption (see below) cannot determine how a particular
voter cast his/her ballot by hashing all ciphertexts in the proof and matching those hashes to the
publicly-posted encrypted vote hashes.

Finally, votes are decrypted and tallied. At this time, Helios also publishes the information
necessary for anyone who wishes to audit the election to verify that the mixnet's shuffling and
decryption were performed properly.

In order to provide a guarantee of proper shuffling, Helios actually shuffles the votes multiple
times, actually using only the final shuffle in tallying the results. The other shuffles are
referred to as ``shadow mixes'' and are made public. Those mixes are hashed to obtain a bit string
that is used, bit-for-bit, as a set of challenges, with one bit value indicating that Helios must
provide the permutation used for the shuffle and the $s_{i}$ values used for re-encryption, and the 
other bit value indiciating that it must provide a mapping from one shadow mix to another. This is
known as an ``Honest-Verifier Zero-Knowlege'' proof, for which the probability that the final
product is correct is equal to $1 - 2^{t}$, where $t$ is the number of shadow mixes produced.

Similarly, the Chaum-Peterson zero-knowledge proof technique is used to prove that decryption is
correct. \todo{How much detail should we provide?}

@@

We will analyze these voting systems by considering each security goal in turn.

@@

In both Scantegrity and Helios, the following information is revealed after the election:
\begin{itemize}
	\item
		\emph{Some} information about \emph{all} encrypted ballots, linked to voters.
	\item
		\emph{All} decrypted ballots, \emph{not} linked to voters.
\end{itemize}

In Scantegrity, ``public'' version of a ballot is really just an encryption of the voter's
selections: it is an independent, uniformly random code chosen beforehand and corresponding to a
particular candidate for that ballot. Table P defines an explicit, ``keyless'' encryption
function -- and each ballot holds just enough information from Table P to encrypt a single voter's
responses. (Alternatively, one can think of Table P as an encryption scheme definition wherein the
ballot number is the key. However, the table is a collection of \emph{all the keys}, and its
compromise would eliminate ballot secrecy for all ballots.)

@@

@@

In Scantegrity, the ballot ``mix'' is predetermined, as it is defined by the generation of Table R.
When that table is constructed, ballots are pseudorandomly ordered. With the exception of audit
ballots, the information revealed that is ultimately revealed allows half of the ballots to be
traced only from Table Q to Table R; and the other half can only be traced from Table R to Table S.
Should Table P be compromised, one could complete all mappings and violate secrecy entirely -- or
simply decrypt encrypted votes with no attention giving to Table R. Should the mapping on
\emph{either} size of Table R (to votes in Table Q or tallies in Table S) be deterministic, privacy
would be compromised for half of all voters.

@@

At the time a ballot is cast, both Scantegrity and Helios attempt to ensure a voter that his/her
vote is recorded exactly as intended.

In Scantegrity, this verification is rolled into the ballot casting process. After recording the
receipt him/herself, a voter can visually check that the code on the receipt matches the candidates
selected on the ballot.

Helios does not actually permit a voter to verify the ballot that he/she actually casts. Rather, it
allows users to verify a ballot without casting it (after which point it cannot be cast without
a new encryption). The assumption here is that \emph{some} voter will choose to audit a ballot, and
so the system cannot simply cheat and show the user a predetermined hash: it \emph{must} be a valid
one -- otherwise the deception would be revealed should the voter choose to exercise his/her right
to audit.

While both systems provide the voter some assurance that the proper vote was recorded, Scantegrity's
verification process is more direct and provides a stronger guarantee that the ballot \emph{actually
cast} was recorded properly.

\subsection{Verifiable collection}

A common theme in verifiable voting systems seems to be the presence of a published ``bulletin
board'' of votes. (Note that because these boards contain only encrypted information, they do not
violate ballot secrecy.)

In the case on Scantegrity, this information is primarily embodied in the published version of Table
Q. After the election, a voter can compare his/her receipt code with that revealed for his ballot in
Table Q. If the code does not match, then something is amiss.

Helios, similarly, exposes a complete list of ballot hashes. A voter can examine this data set to
ensure that it accurately reflects the hash that was reported on screen when the ballot was cast.

@@

\todo{Fill in.}

@@

Scantegrity prevents coercion via the manner it uses to ``encrypt'' ballots. Once cast, a ballot's
ciphertext is \emph{always public}. A voter can claim that he/she voted in a particular manner, but
without exposure of the data maintained in Table P -- Scantegrity's ``secret'' -- the voter has no
way to prove to a third party that the claim is true.

An important point regarding Scantegrity is that it also relies on a \emph{physical property} -- the
invisible ink used when printing confirmation code ``bubbles'' -- to combat vote coercion (i.e., no
one can force a voter to produce a particular receipt, and no one can take a receipt and ably pay
the person for having voted a certain way). This is akin to keeping Table P a secret: the ballots
together \emph{are} Table P, and using special ink allows the codes to be printed without making it
easy for a ballot handler to simple read Table P from the ballots prior to the election.

@@

\subsection{Clever attacks on verifiable voting systems}

\subsection{Voter involvement}

@@

\todo{Remark on the fact that the election itself has a single primary ElGamal keypair.}
